---
title: "Tutorial 09"
author: "Andrew Chou"
date: "27 Oct 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Covid case counts in Singapore
```{r}
library(tidyverse)
library(lubridate)

covid <- read.csv("../data/covid_sg.csv") %>%
  mutate(Date = dmy(Date))
cl <- covid %>%
  pivot_longer(cols = "Quarantined" : "Unlinked", names_to = "status", values_to = "cases") %>%
  group_by(Date) %>%
  mutate(total_cases = sum(cases))

l1 <- "Local Linked and Already Quarantined"
l2 <- "Local Linked and Detected through Surveillance"
l3 <- "Local Unlinked"

ggplot(cl) +
  geom_col(aes(Date, cases, fill = status)) +
  xlab("Press Release Date") +
  ylab("No. of Cases") +
  scale_fill_manual(labels = c(l1, l2, l3), values = c("#006400", "#6CA6C7", "#FFA500")) +
  labs(caption = "Source: Ministry of Health",
       title = "Summary of Local Covid-19 Situation",
       subtitle = "Number of Community Linked and Unlinked Cases") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        legend.justification = "left",
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

## Question 2: Covid cases and deaths around the world.
```{r}
clean_covid_world <- read.csv("../data/covid_world_today.csv")
without_vanuatu <- filter(clean_covid_world, location != "Vanuatu")
vanuatu_txt <- data.frame(x = 100, y = 2, labl = "Vanuatu", continent = "Oceania")

ggplot(clean_covid_world, aes(total_cases, total_deaths, color = continent)) +
  geom_point(alpha = 0.4) +
  facet_wrap(~continent) +
  scale_x_log10() +
  scale_y_log10() +
  geom_text(data = vanuatu_txt, aes(x, y, label = labl), show.legend = FALSE) +
  geom_smooth(data = without_vanuatu, method = "lm", se = FALSE, lty = "dashed", color = "black") +
  xlab("Total Cases") +
  ylab("Total Deaths")
```

## Question 3: Olympic athletes data
```{r, warning = FALSE}
noc <- read.csv("../data/athlete_events.csv")
library(GGally)
noc %>%
  filter(Year >= 1970, Season == "Summer", Sport %in% c("Athletics", "Swimming")) %>%
  group_by(ID) %>%
  summarise(Sex = Sex[1], Age = min(Age, na.rm = FALSE),
            Height = mean(Height, na.rm = FALSE),
            Weight = mean(Weight, na.rm = FALSE), Sport = Sport[1], Medal = any(!is.na(Medal)),
            .groups = "drop") -> noc_tmp
ggpairs(noc_tmp, mapping = aes(color = Sport, alpha = 0.8), columns = 2:7,
        lower = list(continuous = "density", discrete = "crosstable"),
        showStrips = TRUE)
```

### My observations:

1. The median age of swimmers is much lower than that of athletes.
2. The median height of medal winners is slightly more than that of non-winners.

### What I would explore next:

1. Has the median age of swimmers always been less than that of athletes? I will plot the median age of swimmers and athletes by year using the original data set.

2. The median height of all medal winners is slightly more than that of non-winners. But is this true for every sport? Which sport has the greatest height advantage? Is there any sport with a height disadvantage? I will compare the median height of medal winners and non-winners in each sport.