---
title: "Tutorial 5"
author: "Andrew Chou"
date: "23 Spetember 2021"
output:
  html_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Number of couriers
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
absent <- read_csv2("../data/Absenteeism_at_work.csv") %>% mutate(ID = as.integer(ID))
couriers <- absent %>% select("ID") %>% n_distinct()
couriers
```
There are `r couriers` couriers in the dataset.

## Question 2: Max absences
```{r}
max_absence_tbl <- absent %>% select(ID, `Reason for absence`, `Absenteeism time in hours`) %>% top_n(n = 7)
max_absence_tbl
```

## Question 3: Unknown reason for absence
```{r}
reason_tbl <- absent %>% select(ID, `Reason for absence`, `Disciplinary failure`, `Absenteeism time in hours`) %>% filter(`Reason for absence` == 0)
reason_tbl
```
I observe that all but three of the couriers have disciplinary failure, and the absenteeism time are all 0 hours.

## Question 4: Remove disciplinary failure
```{r}
absent <- absent %>% filter(`Disciplinary failure` == 0)
dim(absent)
```

## Question 5: Recode day of week and season
```{r}
absent <- absent %>% mutate(`Day of the week` = recode(`Day of the week`, `2` = "Mon", `3` = "Tue", `4` = "Wed", `5` = "Thu", `6` = "Fri")) %>% mutate(Seasons = recode(Seasons, `1` = "Summer", `2` = "Autumn", `3` = "Winter", `4` = "Spring"))
absent %>% select(ID, `Day of the week`, Seasons) %>% sample_n(10)
```

## Question 6: Proportion of absences in Day by Season
```{r}
rows <- nrow(absent)
seasons_tbl <- absent %>% group_by(Seasons) %>% count()

absence_props <- absent %>% group_by(Seasons, `Day of the week`) %>% count() %>% full_join(seasons_tbl, by = "Seasons") %>% mutate(n = n.x, prop = n.x / n.y) %>% select(-c(n.x, n.y))

absence_props
```

## Question 7: Summaries by courier
```{r}
absent_stats <- absent %>% select(ID, `Absenteeism time in hours`) %>% group_by(ID) %>% summarize(abs_time = sum(`Absenteeism time in hours`), abs_count = n())

courier_summary <- absent %>% select(ID, `Absenteeism time in hours`) %>% group_by(ID) %>% summarize(v = quantile(`Absenteeism time in hours`), q = c("min", "lower", "median", "upper", "max")) %>% pivot_wider(names_from = q, values_from = v) %>% inner_join(absent_stats) %>% mutate(abs_time = as.integer(abs_time))

courier_summary
```
The top 3 diligent couriers have ID numbers 4, 8, and 35. The bottom 3 are numbers 3, 14, and 11.

## Question 8: Status Changes
```{r}
absent %>% select(ID, Education:Pet) %>% distinct() %>% dim()
```
The output tibble has 36 rows, one for each courier. This tells us that none of the couriers’ status changed over
the three years. Even their weight didn't change!
